



%__________   ACKNOWLEDGMENTS  ______________________________
\vspace*{0.5in}
\begin{center}
\section*{ACKNOWLEDGMENTS}
\end{center}
\addcontentsline{toc}{section}{ACKNOWLEDGMENTS}


\noindent
This study was supported by the Institutional Development Award (IDeA) INBRE Grant Number P20GM103475 from the National Institute of General Medical Sciences (NIGMS), a component of the National Institutes of Health (NIH) and the Bioinformatics Research Core of INBRE. Its contents are solely the responsibility of the authors and do not necessarily represent the official view of NIGMS or NIH.

I would like to express my deepest gratitude to my wonderful wife Nichool; your patience, understanding, and countless sacrifices made it possible for me to focus on my research. Your faith in my abilities has been my rock. To my family Edwin, Angy and Salomé; for their unwavering support, love, and encouragement throughout this academic journey. Your belief in me has been my constant source of motivation. I am also profoundly grateful to my esteemed professor Dr. Almodóvar; for his mentorship, guidance, and expertise. Your insights and dedication to my academic growth have been invaluable.

This thesis is a culmination of not just my efforts, but the collective support and faith of my loved ones. I am thankful beyond words for your presence in my life. 
% PASTE YOUR ACKNOWLEDGMENTS HERE (DELETE \blindtext)
%____________________________________________________________


\newpage



%__________   DEDICATION  ______________________________
\vspace*{2in}
\begin{center}
\emph{This is for Sócrates and/or Luna} %USE THIS SPACE FOR DEDICATION (IF NOT, DELETE)
\end{center}
%____________________________________________________________
